---
title: "Basic R"
output:
  html_document:
    df_print: paged
---

# First steps

```{r someSetup}
# clean memory
rm(list = ls())
```


# Collecting Data}

```{r collecting_DataTable}
# github link to data:
link="https://github.com/eScienceWinterSchool/R_Session/raw/main/data/hdidemiso_plus.RDS"

#opening data
fileLink=url(link)
mergedData=readRDS(fileLink)
```

knowing column names

```{r knowing column names}
names(mergedData)
```

data types

```{r data types}
str(mergedData)
```


# categorical exploration 

## frequency table

```{r frequency table}
library(summarytools)

freq(mergedData$Regimetype, 
     plain.ascii = FALSE,
     report.nas=FALSE)
```

## compute statistics

```{r compute statistics}
DescTools::Mode(mergedData$Regimetype)
DescTools::Median(mergedData$Regimetype)

```


# Plots
        
```{r prepare plot}

library(ggplot2)

base=ggplot(data = mergedData)
bar=base + geom_bar(aes(x=Regimetype),
                    fill='orange',
                    color='red')
bar + labs(title = "Democracy in the world (2021)",
           x= "Regime Type",
           y= "count of countries",
           caption = 'Source: The Economist')

```


# Numeric exploration 

## summary statistics

```{r table of summary statistics}
summary(Filter(is.numeric, mergedData))
```

The Boxplot

```{r numBoxplot}
mM_Positions=grep('mM',names(mergedData))
DataForBox=reshape2::melt(mergedData[,mM_Positions])

base=ggplot(data=DataForBox)
box=base + geom_boxplot(aes(x=variable,y=value),
                        fill='lightblue',
                        color='red')
box + coord_flip()
```
        
        

# Linear regression 

Hipotheses and regressions

```{r Hipotheses and regressions}
h1=formula(HumanDevelopmentIndex~Functioningofgovernment)
regre1=glm(h1,data=mergedData,family = "gaussian")

h2=formula(HumanDevelopmentIndex~Functioningofgovernment + Politicalparticipation)
regre2=glm(h2,data=mergedData,family = "gaussian")
```

## Results

### first regression
```{r show result of first regression}
summary(regre1)
library(rsq)
rsq(regre1)
```
### second regression

```{r show result of second regression}
summary(regre2)
lm.beta::lm.beta(regre2)
rsq(regre2,adj=T)
```

### compare regressions

```{r compare regressions}
anova(regre1,regre2,test = 'Chisq') 
```


# Logistic regression 


## Hipotheses and regressions

```{r Hipotheses and regressions logit}
# data selection
Selection=c('Authoritarian','Flawed democracy')
mergedData_Dico=mergedData[mergedData$Regimetype%in%Selection,]

h1=formula(Regimetype~MeanYearsOfSchooling)
regreLogi1=glm(h1,data=mergedData_Dico,family = "binomial")

h2=formula(Regimetype~MeanYearsOfSchooling + LifeExpectancyAtBirth)
regreLogi2=glm(h2,data=mergedData_Dico,family = "binomial")
```

## Results

### first regression

```{r show result of first regression logit}
summary(regreLogi1)

```

### second regression

```{r show result of second regression logit}
summary(regreLogi2)
```

### compare regressions

```{r compare regressions logit}
anova(regreLogi1,regreLogi2,test = 'LRT')
```

# Clustering 

## subsetting

```{r subset}

positions_mM=grep("mM", names(mergedData)) #easy way detect positions
dataToCluster= mergedData[,positions_mM]
row.names(dataToCluster)=mergedData$Countryname

```


## computing clusters

```{r clustering with k-medoids}
set.seed(111)
library(cluster)
distances=cluster::daisy(dataToCluster)
res.pam=pam(distances,4,cluster.only = F)
dataToCluster$pam=res.pam$cluster

```

## understanding labels

```{r understanding cluster labels}
pamMeans=aggregate(data=dataToCluster,Overallscore_mM~pam,mean)
pamMeans[order(pamMeans$Overallscore_mM),]
```

## recoding labels

```{r recoding labels}
#library(dplyr)
recoded=dplyr::recode_factor(dataToCluster$pam,
                             `3`=1,`1`=2,`2`=3,`4`=4,
                             .ordered = T)

# new column 
mergedData$pam=recoded
```
                                       

# Mapping

The map data:

```{r open file}
# link in Github
linkMap="https://github.com/eScienceWinterSchool/R_Session/raw/main/worldMap.geojson"
# read in file
library(sf)
mapWorld=read_sf(linkMap)
```


```{r columns}
names(mapWorld)
```


```{r quick look}
head(mapWorld)
```


```{r structure}

str(mapWorld)
```

## merging variables into map

```{r merge}

mapWorld_Data=merge(mapWorld,mergedData,
                    by.x = 'ISO3', 
                    by.y = 'iso3')

```

## Plotting

```{r plot}
baseMap=ggplot(mapWorld) + 
     geom_sf(color=NA,fill='grey90') + 
     theme_light()

mapPlot=baseMap + geom_sf(data = mapWorld_Data,
                       aes(fill=pam),
                       size=0.1,
                       color='grey50') + 
        labs(title = "Countries by HDI & Democracy",
             subtitle = "(missing values in gray)",
             caption = "Source: The Economist") 

mapPlot + scale_fill_brewer(palette = "YlOrRd",
                            name='Levels (ascending)')

```



